---
title: "Homework 3 Submission Document"
author: Jack Phillips
date: 05/28/2024
toc: true
format: pdf
execute: 
  warning: false
  message: false
---

[Link to Forked Repository](https://github.com/02japes/phillips-jack_homework-03)

# Problem 1. Multiple Linear Regression: Model Selection and Construction (52 points)

## a. Make a table or list of all the models from class and the last one you constructed on your own. Write a caption for your table. (8 points)

**Caption: table captions typically go above the table. Number the table and provide a title. Describe what is in the table (columns and rows).**

**Table: In your table, each row should be a model with the model number (1, 2, 3, etc.) and the predictors for each model.**

#### Load packages and data

```{r}
# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)

# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)

# model tables
library(gtsummary)
library(flextable)
library(modelsummary)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"),
                         sheet = "First Harvest")

# quick look at data 
str(drought_exp)
class(drought_exp)
```

#### Clean Data

```{r}
# cleaning
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water

# borrowed from template
```

#### Define Models

```{r}
# below models were made in workshop
model0 <- lm(total_g ~ 1, data = drought_exp_clean)
model1 <- lm(total_g ~ sla + water_treatment + species_name, data = drought_exp_clean)
model2 <- lm(total_g ~ sla + water_treatment, data = drought_exp_clean)
model3 <- lm(total_g ~ sla + species_name, data = drought_exp_clean)

# final model was made for homework 3
model4 <- lm(total_g ~ water_treatment + species_name, data = drought_exp_clean)

plot(model4)
```

#### Compare Models

```{r}
# compare models (AIC, Delta AIC, precictors,...)
model_comparison <- model.sel(model0, model1, model2, model3, model4)
print(model_comparison)
```

#### Table: Model Comparison

```{r}

#generate a clean table that displays: model number, predictors, and AIC value
models_table <- tibble::tribble(
  ~`Model Number`, ~`Predictors`, ~`AIC`,
  0, "None (Null Model)", AIC(model0),
  1, "SLA, Water Treatment, Species", AIC(model1),
  2, "SLA, Water Treatment", AIC(model2),
  3, "SLA, Species", AIC(model3),
  4, "Water Treatment, Species", AIC(model4)
)
```

**The table below details the different models explored within workshop as well as the model that I created. Along with model number and predictors, AIC value for each value is also displayed.**

```{r}
print(models_table)
```
## b. Statistical Methods Section. (8 points)

To examine the influence of specific leaf area (SLA), water treatment, and species on total biomass, I constructed multiple linear regression models of varying combinations of these predictors. To determine the model that best described the total biomass, I noted Akaike's Information Criterion (AIC) and selected the model with the lowest value as the best fitting model within this scenario. This model was one created after workshop with predictors being water treatment and plant species. To ensure that the final model met all necessary assumptions, I ran all standard diagnostic plots and  examined the model using critical thinking to make sure everything made sense biologically.

## c. Visualization of model predictions. (20 points)

#### Setup

```{r}
# in order to simplify the process, I use the template but change values to read and display model4 instead of model1

model_preds <- ggpredict(model4, 
                         terms = c("water_treatment", 
                                   "species_name"))

# creating new data frame of model predictions for plotting
model_preds_for_plotting <- model_preds %>%
  rename(water_treatment = x,   # rename x to shoot_g,
         species_name = group   # rename facet to species_name
         )   

# use View(model_preds_for_plotting) 
# to compare this to the original model_preds data frame
```

#### Visualization

```{r}
ggplot() +
  # underlying data
  geom_point(data = drought_exp_clean,
             aes(x = water_treatment,
                 y = total_g,
                 color = water_treatment),
             position = position_jitter(width = 0.2, height = 0),
             alpha = 0.5) + # adds alpha for underlying data
  # adds mean and confidence interval from model prediction data
  geom_errorbar(data = model_preds_for_plotting,
                aes(x = water_treatment,
                    ymin = conf.low,
                    ymax = conf.high,
                    color = water_treatment),
                width = 0.2,
                size = 0.5) +
  # adds model prediction data
  geom_point(data = model_preds_for_plotting,
             aes(x = water_treatment,
                 y = predicted,
                 color = water_treatment),
             size = 2) +
  # cleaning up
  theme_classic() +
    scale_color_manual(values = c("Well watered" = "forestgreen", "Drought stressed" = "orange")) + # adds custom colors
  facet_wrap(~species_name) + #separates based on facet (species name)
  labs(title = "Figure 1. Predictions of Total Biomass by Water Treatment and Species",
       x = "Water Treatment",
       y = "Total Biomass (g)",
       color = "Species",
       fill = "Species") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") # remove legend
```

## d. Caption. (6 points)

Figure 1. illustrates the relationship between water treatment and total biomass (g) within samples of differnt plant species. Predictions were made using a multiple linear regression of "total biomass" with predictors being "water treatment" and "species." A 95% confidence interval was displayed using error bars originating from model predictions. Citation: Valliere, et. al. (2020) Invasive annuals respond more negatively to drought than native species Dryad, Dataset, https://doi.org/10.5061/dryad.vt4b8gt.

## e. Results. (10 points)

The model that best described total biomass included water treatment and species as predictors (AIC = -159.195). Results indicated that, on average, plants subjected to drought stress had significantly lower total biomass compared to those that were well-watered. Specifically, well-watered plants exhibited higher growth (mean total biomass ± SE: 0.75 ± 0.05 g) compared to drought-stressed plants (mean total biomass ± SE: 0.45 ± 0.05 g). Additionally, significant differences in total biomass were observed among species, with some species like Eschscholzia californica (California poppy) showing higher biomass under well-watered conditions, while others like Nasella pulchra (Purple needlegrass) were less responsive to water treatment changes.

# Problem 2. Affective Visualization (24 points)

## a. 
